\magnification=\magstep1
\nopagenumbers

\centerline{Special University Oral Examination}

\bigskip

\centerline{\bf A Computational Investigation of Receptivity in}
\centerline{\bf High-Speed Flow Near a Swept Leading-Edge}

\bigskip

\centerline{S. Scott Collis}
\centerline{Mechanical Engineering Department}
\centerline{Stanford University}

\bigskip

\centerline{3:00 PM Thursday, November 7}
\centerline{Durand Building, Room 450}
\smallskip
\centerline{\it Refreshments served at 2:45 PM}

\bigskip
\bigskip

Laminar-flow wings have the promise of reducing viscous drag forces in cruise
for commercial aircraft.  However, the success of a laminar-flow wing depends
critically on the external disturbance environment and how these disturbances
influence the transition from laminar to turbulent flow.  The process by which external disturbances are converted into instability waves, which are the precursors to turbulence, is called receptivity.
\smallskip
The research described here focuses on the receptivity of the three-dimensional boundary layer near the leading edge of a high-speed swept-wing.  In particular, the influence of surface roughness near the leading edge is examined as it relates to the formation of stationary crossflow vortices in the boundary layer.  Recent experiments indicate that surface roughness is the primary cause of stationary crossflow vortices which are observed to dominate the laminar to turbulent transition process on swept wings.  Therefore, the mechanisms responsible for the formation of crossflow vortices and the accurate prediction of their initial amplitude are essential for the development of laminar-flow wings.
\smallskip
The receptivity and early evolution of stationary crossflow vortices is investigated through numerical solutions of the Navier--Stokes equations for a swept leading-edge.  Consistent with findings for other geometries, convex surface curvature stabilizes crossflow vortex growth while nonparallel effects are destabilizing.  In contrast, the initial amplitude of crossflow vortices downstream of a localized surface roughness site is found to be greater in the presence of convex surface curvature, while the nonparallel meanflow near a leading-edge strongly reduces the initial amplitude of crossflow vortices.  These competing effects---curvature and nonparallelism---tend to counteract one another, but, for the conditions studied here, the nonparallel effect is dominant.
\smallskip
Comparisons between Navier--Stokes solutions and recent theoretical receptivity analysis, based on the parallel-flow equations, show that the theoretical method over-predicts the initial amplitude of stationary crossflow vortices by as much as $77\%$ for long wavelength disturbances.  The error in the theoretical prediction is reduced for shorter wavelengths, since nonparallel effects are relatively less important, but remains as high as $30\%$ near the leading-edge.  Although the current theory provides a conservative estimate of the initial amplitude of crossflow vortices, it is concluded that accurate theoretical prediction of crossflow receptivity for realistic geometries requires the inclusion of nonparallel effects.  

\end
