%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  Acknowledgements
%
%  S. Scott Collis
%
%  Written: 9-5-95
%
%  Revised: 10-29-96
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

I thank my advisor, Professor Sanjiva K. Lele, for his excellent technical
advice and the fundamental background he has provided me through the courses
he has taught.  His experience and wisdom is greatly appreciated and I look
forward to further collaborations.

I am also indebted to Professors Parviz Moin and William C. Reynolds for
serving on my dissertation reading committee, and Professors Robert
W. MacCormack and Stephen G. Monismith who served on my University examination
committee.

Many people have contributed to my education at Stanford both in and out of
the classroom.  Certainly a large impact has come from my friends and
office-mates: Ganyu Lu, Krishnan Mahesh, Keith Lucas, Steve Rennich, Ted
Manning, Sungwon Bae, and Calvin Lui.  Through countless conversations, they
have offered their help and assistance and it is much appreciated.

Finally, I thank my parents for their constant support throughout my studies,
and, for her understanding, encouragement and love, I dedicate this work to my
wife, Emily.

\vskip 1.0in

I was awarded a National Science Foundation Fellowship that supported the
first three years of this research.  Subsequent funding was supplied by the
National Science Foundation through a P.Y.I. award to Professor Sanjiva Lele
with industrial matching funds from the Boeing Company.  Computer time was
provided by NAS through the NASA Ames Research Center and by the National
Science Foundation through the San Diego Supercomputer Center.

