%===============================================================================
%
%  Abstract for University Microfilms
%
%  Note:  Must be less than 350 words
%
%  Author: Scott Collis
%
%  Revised: 3-12-97
%
%===============================================================================
%\input lcdplain
\magnification=\magstep1
\baselineskip= 16.5pt plus .1pt

\centerline{\uppercase{A Computational Investigation of Receptivity}}
\centerline{\uppercase{in High-Speed Flow Near a Swept Leading-Edge}}
\bigskip
\centerline{ S. Scott Collis, Ph.D.}
\centerline{Stanford University, 1997}
\bigskip
\centerline{ Adviser: Sanjiva K. Lele, Ph.D.}
\bigskip

Laminar-flow wings show promise for reducing viscous drag forces for
commercial aircraft.  However, the success of a laminar-flow wing depends
critically on the external disturbance environment and how these disturbances
influence the transition from laminar to turbulent flow.  The process by which
external disturbances are converted into instability waves, which are the
precursors to turbulence, is called receptivity.

The research described in this thesis focuses on the receptivity of the
three-dimensional boundary-layer near the leading edge of a high-speed swept
wing.  In particular, the influence of surface roughness is examined as it
relates to the formation of stationary crossflow vortices in the boundary
layer.  Recent experiments indicate that stationary crossflow vortices,
initiated by surface roughness, dominate the laminar-to-turbulent transition
process on swept wings.  Therefore, the mechanisms responsible for their
formation and the accurate prediction of their initial amplitude are essential
for the development of laminar-flow wings.

The receptivity and early evolution of stationary crossflow vortices is
investigated through numerical solutions of the linearized Navier--Stokes
equations for a swept leading-edge.  Consistent with findings for other
geometries, convex surface curvature stabilizes crossflow vortex growth while
nonparallel effects are destabilizing.  In contrast, the initial amplitude of
crossflow vortices downstream of a localized surface roughness site is greater
in the presence of convex surface curvature, while the nonparallel meanflow
near a leading-edge strongly reduces the initial amplitude of crossflow
vortices.  These competing effects tend to counteract one another, but, for
the conditions studied here, the nonparallel effect is dominant.

Comparisons between linearized Navier--Stokes solutions and recent theoretical
receptivity analysis, based on the parallel-flow equations, show that the
theoretical method over-predicts the initial amplitude of stationary crossflow
vortices by as much as $77\%$ for long wavelength disturbances.  The error in
the theoretical prediction is reduced for shorter wavelengths, since
nonparallel effects are relatively less important, but remains as high as
$30\%$ near the leading-edge.  Although parallel theory provides a
conservative estimate of the initial amplitude of crossflow vortices, it is
concluded that accurate prediction of crossflow receptivity near a swept
leading-edge requires the inclusion of nonparallel effects.

\bigskip
\bigskip
\vbox{\hbox to \hsize{\hfil\vbox{
\settabs\+By:\ &For the Department of Mechanical Engineering\cr
\+Approved for publication:&\cr
\+&\cr
\+By:&\underbar{\hbox{\hskip 2.75in}}\cr
\+&For the Department of Mechanical Engineering\cr}}}

\bye
