%========================================================================
%
%  LNS3D Notes
%
%  S. Collis
%
%  3-14-2001
%========================================================================
\documentclass[12pt]{article}
\usepackage{mathtime}
\usepackage{fullpage}
\usepackage{cite}
\usepackage{xspace}
\usepackage{amsbsy}
\usepackage{subeqn}
\newif\ifCUPmtlplainloaded
\usepackage{upmath}
\input{nepsf}
%\input{macros}

\begin{document}

\title{Notes on LNS3D}
\author{S. Scott Collis}
\maketitle

\section*{03-14-2001}

\begin{enumerate}

\item Implemented extrapolation of $\rho$ in the nonlinear solver at the wall
for both explicit and implicit time-advancement ({\tt wall=3}).  It works fine
in both solvers but it is much better than continuity at the wall when using
the explicit solver.  In the implicit solver, the convergence is not quite as
rapid when using extrapolation, especially in the first iteration.

\item I also put the ({\tt wall=3}) condition in the second-order LHS.

\item The explicit solver is stable for $CFL<=2.0$ when using $4^{th}$ order
spatial differences.  The $6^{th}$ order requires $CFL<=1.5$.

\item Right now I have $\epsilon_e=0.125$ which is rather high.  I will try
reducing it to $0.0125$.

\item I found an error in {\tt genmtrx.f90}.  Basically I wasn't putting the
$D$ matrix into $\hat D$ thereby leaving $\hat D$ either zero or undefined!
This may be what was causing Nawaf's difficulties.  It wouldn't make the
nonlinear code incorrect, although it would affect convergence.  It would,
however, make the linear code incorrect!

\item The Reimann boundary conditions on the LHS are not supported for the
$s^{nd}$ order LHS right now -- I need to change the call to Reimann LHS to
reflect the change in the data structure at the boundary.  No, this wasn't it.
Instead I need to pass local values of rho, ub, vb, etc. on the boundary which
come from the potential solution.  Now it appears to be working correctly.

\item Switched back to the 3d test and still something is wrong...  There is a
big change from the IC to the first time-step.

\end{enumerate}

\section*{03-19-2001}

\begin{enumerate}

\item The implicit solver saturates at the wall corners on the vortex pair
test case since both the Reimann and wall boundary conditions are applied
here!

\item Added Lele--Poinsot to genmtrx (nonlinear) and nrk for wall=4.  Need
to revised the linearized LHS.

\item Tried out the L\&P BC on the viscous vortex rebound problem and it works
well.

\end{enumerate}

\section*{03-21-2001}

\begin{enumerate}

\item Worked on AR-Pack eigenvalue solver.  Got the shift and invert to work
the the acoustic wave model problem using the approximate factorization
iterative solver.

\item Approximate factorization seems to be very slow.

\item I put in a point Jacobi iteration and it doesn't converge!

\end{enumerate}

\section*{03-21-2001}

\begin{enumerate}

\item Updated comments in si\_eigdrv3d.f90.

\end{enumerate}

\end{document}
